---
title: "European Revision"
output: html_notebook
---

Checking out the scatterplot data of the GEZ and forest.  I took a random sampling of points, 2000 points per zone.  I then extracted the data for:

1. [Mean annual temperature](https://developers.google.com/earth-engine/datasets/catalog/WORLDCLIM_V1_BIO)
2. [Elevation](https://developers.google.com/earth-engine/datasets/catalog/USGS_GMTED2010)
3. [SPEI](https://developers.google.com/earth-engine/datasets/catalog/ECMWF_ERA5_MONTHLY)

I calculated the SPEI using the [SPEI package](https://cran.r-project.org/web/packages/SPEI/index.html) in R and calculated the mean SPEI for the monthly time period between 2000 - 2020.

For the plots below, *pts* represent the random points taken for each GEZ and *sites* represent the primary forest sites in the paper.  The plots are interactive and you can zoom in and out and move axes of the plot, just click on the plot.

```{r libraries, echo = F, results='hide', message=FALSE, error = F, include = F, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(scatterplot3d)
library(FactoClass)
# library(plotly)
```

```{r formatData, echo = F, results='hide', message=FALSE, error = F, include = F, warning=FALSE}

eurPts<-read.csv('E:\\Big_Tree\\data\\Europe\\revision\\europeanPointsExtract.csv')
eurSites<-read.csv('E:\\Big_Tree\\data\\Europe\\revision\\europeanSitesExtract.csv')
eurPts$gezValue<-gsub("\\[|\\]","", eurPts$gezValue)
eurSites$gezValue<-gsub("\\[|\\]","", eurSites$gezValue)

#------------------------------------------------------------------------
#sites
#------------------------------------------------------------------------
sites<-data.frame(as.numeric(eurSites$ID), 
                  as.factor(eurSites$Site_Type),
                  as.character(eurSites$gezValue),
                  as.numeric(eurSites$meanSpeiValue),
                  as.numeric(eurSites$minSpeiValue),
                  as.numeric(eurSites$meanTempValue),
                  # as.numeric(eurSitesBioclim$meanTempValue*0.1),
                  as.numeric(eurSites$elevation/100),
                  as.numeric(eurSites$forestCover30Value))

head(sites)
names(sites)<- c('ID', 'type', 'gez', 'meanSpei', 'minSpei', 'meanTemp', 'elevation', 'forest30')

#add rounded data
sites$meanSpeiRd <- round(sites$meanSpei, 2)
sites$minSpeiRd <- round(sites$minSpei, 2)

#omit all n/a rows
sites<-na.omit(sites)

sites30 <- dplyr::filter(sites, forest30 %in% c(1))
research<- dplyr::filter(sites, type %in% c('research'))
head(sites30)

#------------------------------------------------------------------------
#points
#------------------------------------------------------------------------
eurPts$type <- "pts"
pts<-data.frame(as.numeric(eurPts$CID + 6870),
                as.factor(eurPts$type),
                as.character(eurPts$gezValue), 
                as.numeric(eurPts$meanSpeiValue), 
                as.numeric(eurPts$minSpeiValue), 
                as.numeric(eurPts$meanTempValue),  
                as.numeric(eurPts$elevation/100),
                as.numeric(eurPts$forestCover30Value))
head(pts)
names(pts)<- c('ID', 'type', 'gez', 'meanSpei', 'minSpei', 'meanTemp', 'elevation', 'forest30')

#add rounded data
pts$meanSpeiRd <- round(pts$meanSpei, 2)
pts$minSpeiRd <- round(pts$minSpei, 2)

#omit all n/a rows
pts<-na.omit(pts)

pts30 <- dplyr::filter(pts, forest30 %in% c(1))


head(pts30)

#-------------------------
combine<-rbind(pts30, sites30)
# research<-dplyr::filter(combine, type %in% c('research'))
```

```{r plotLimits,echo=FALSE, message=FALSE, warning=FALSE}
max_x <- max(combine$meanTemp) #mean Annual temp
min_x <- min(combine$meanTemp)
max_y <- max(combine$meanSpei) #mean SPEI value
min_y <- min(combine$meanSpei)
max_z <- max(combine$elevation) #mean Annual temp
min_z <- min(combine$elevation)

```

```{r borealConiferous, echo=FALSE, message=FALSE, warning=FALSE}

#Look at 3d scatterplot
gezType <-'Boreal coniferous forest'
combinePts<-dplyr::filter(combine,gez == gezType )
research<-dplyr::filter(combinePts, type %in% c('research'))
# colourCombo <- c("#98FB9825", "#00FFFF75", "#FFFF6675", "#FF573375")
colourCombo <- c("#B1E5AD30", "#00CCCC", "#F6D65890", "#C9242C90")
# colourCombo<- c("#76C56E25","#008B8B75", "#DAA52090", "#C9242C90")
# Pale Green: #98FB98 pts
# Red: #FF5733 research sites
# Aqua: #00FFFF literature sites
# Yellow: #FFFF66 inventory sites
colours<-colourCombo[as.numeric(combinePts$type)]
# scatterplot3d(x = combinePts$meanTemp, 
#               y = combinePts$meanSpei, 
#               z = combinePts$elevation,
#               pch = 20,
#               color = colours,
#               angle = 70,
#               main = gezType,
#               xlab = "Mean annual Temperature(°C)",
#               ylab = "Water availability index",
#               zlab = "Elevation (100m)",
#               xlim=c(min_x,max_x),
#               ylim=c(min_y,max_y),
#               zlim=c(0,max_z)
#               )
# legend()

# source('~/hubiC/Documents/R/function/addgrids3d.r')
# http://www.sthda.com/english/wiki/scatterplot3d-3d-graphics-r-software-and-data-visualization
s3d <- scatterplot3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei, 
              z = combinePts$elevation,
              pch = "",
              grid = F,
              box = T,
              # angle = 70,
              # color = colours,
              main = gezType,
              xlab = "Mean annual Temperature(°C)",
              # ylab = "",
              ylab = "Water availability index",
              zlab = "Elevation (100m)",
              xlim=c(min_x,max_x),
              ylim=c(min_y,max_y),
              zlim=c(0,max_z)
              )

addgrids3d(x = combinePts$meanTemp, 
           y = combinePts$meanSpei, 
           z = combinePts$elevation,
           grid = c("xy", "xz", "yz"),
           # col.grid = "grey",
           xlim=c(min_x,max_x),
           ylim=c(min_y,max_y),
           zlim=c(0,max_z)
           )
s3d$points3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei,
              z = combinePts$elevation,
              pch = 20,
              col= colours)

# dims <- par("usr")
# x <- dims[1]+ 0.9*diff(dims[1:2])
# y <- dims[3]+ 0.08*diff(dims[3:4])
# mtext(x,y,"Water availability index",srt=30)
legend(s3d$xyz.convert(20, -0.5, -4),
       legend = c("forest cover random points",
                  "research sites",
                  "literature sites",
                  "inventory sites"
                  ),
       col = c("#B1E5AD75", "#C9242C90", "#00CCCC", "#F6D65890"),
       pch = 20,
       cex = 0.8,
       inset = 0.00, xpd = TRUE, horiz = F)
```

```{r borealMs, echo=FALSE, message=FALSE, warning=FALSE}

#Look at 3d scatterplot
gezType <-'Boreal mountain system'
# ppPts<-dplyr::filter(pts30, id == 'pts' & gez == gezType )
combinePts<-dplyr::filter(combine, gez == gezType )
# research<-dplyr::filter(combinePts, type %in% c('research'))
colours<-colourCombo[as.numeric(combinePts$type)]
s3d <- scatterplot3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei, 
              z = combinePts$elevation,
              pch = "",
              grid = F,
              box = T,
              # angle = 70,
              # color = colours,
              main = gezType,
              xlab = "Mean annual Temperature(°C)",
              ylab = "Water availability index",
              zlab = "Elevation (100m)",
              xlim=c(min_x,max_x),
              ylim=c(min_y,max_y),
              zlim=c(0,max_z)
              )

addgrids3d(x = combinePts$meanTemp, 
           y = combinePts$meanSpei, 
           z = combinePts$elevation,
           grid = c("xy", "xz", "yz"),
           # col.grid = "grey",
           xlim=c(min_x,max_x),
           ylim=c(min_y,max_y),
           zlim=c(0,max_z)
           )
s3d$points3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei,
              z = combinePts$elevation,
              pch = 20,
              col= colours)

legend(s3d$xyz.convert(20, -0.5, -4),
       legend = c("forest cover random points", 
                  "research sites",
                  "literature sites", 
                  "inventory sites" 
                  ),
       col = c("#B1E5AD60", "#C9242C90", "#00FFFF80", "#F6D65890"),
       pch = 20,
       cex = 0.7,
       inset = 0.00, xpd = TRUE, horiz = F)
```

```{r TemperateContinental Forest, echo=FALSE, message=FALSE, warning=FALSE}

#Look at 3d scatterplot
gezType <-'Temperate continental forest'
# ppPts<-dplyr::filter(pts30, id == 'pts' & gez == gezType )
combinePts<-dplyr::filter(combine, gez == gezType )
# research<-dplyr::filter(combinePts, type %in% c('research'))
colourCombo <- c("#B1E5AD30", "#00CCCC", "#F6D65880", "#C9242C85")
# Pale Green: #B1E5AD30 pts
# Red: #C9242C research sites
# Aqua: "#00FFFF80"literature sites
# Yellow: #F6D65890 inventory sites
colours<-colourCombo[as.numeric(combinePts$type)]
s3d <- scatterplot3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei, 
              z = combinePts$elevation,
              pch = "",
              grid = F,
              box = T,
              # angle = 70,
              # color = colours,
              main = gezType,
              xlab = "Mean annual Temperature(°C)",
              ylab = "Water availability index",
              zlab = "Elevation (100m)",
              xlim=c(min_x,max_x),
              ylim=c(min_y,max_y),
              zlim=c(0,max_z)
              )

addgrids3d(x = combinePts$meanTemp, 
           y = combinePts$meanSpei, 
           z = combinePts$elevation,
           grid = c("xy", "xz", "yz"),
           # col.grid = "grey",
           xlim=c(min_x,max_x),
           ylim=c(min_y,max_y),
           zlim=c(0,max_z)
           )
s3d$points3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei,
              z = combinePts$elevation,
              pch = 20,
              col= colours)

legend(s3d$xyz.convert(20, -0.5, -4),
       legend = c("forest cover random points", 
                  "research sites",
                  "literature sites", 
                  "inventory sites" 
                  ),
       col = c("#B1E5AD60", "#C9242C90", "#00CCCC", "#F6D65890"),
       pch = 20,
       cex = 0.7,
       inset = 0.00, xpd = TRUE, horiz = F)
```

```{r TemperateMountainSystem, echo=FALSE, message=FALSE, warning=FALSE}

#Look at 3d scatterplot
gezType <-'Temperate mountain system'
# ppPts<-dplyr::filter(pts30, id == 'pts' & gez == gezType )
combinePts<-dplyr::filter(combine, gez == gezType )
colourCombo <- c("#B1E5AD30", "#00CCCC", "#F6D65880", "#C9242C85")
# Pale Green: #B1E5AD30 pts
# Red: #C9242C research sites
# Aqua: "#00FFFF80"literature sites
# Yellow: #F6D65890 inventory sites
colours<-colourCombo[as.numeric(combinePts$type)]
s3d <- scatterplot3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei, 
              z = combinePts$elevation,
              pch = "",
              grid = F,
              box = T,
              # angle = 70,
              # color = colours,
              main = gezType,
              xlab = "Mean annual Temperature(°C)",
              ylab = "Water availability index",
              zlab = "Elevation (100m)",
              xlim=c(min_x,max_x),
              ylim=c(min_y,max_y),
              zlim=c(0,max_z)
              )

addgrids3d(x = combinePts$meanTemp, 
           y = combinePts$meanSpei, 
           z = combinePts$elevation,
           grid = c("xy", "xz", "yz"),
           # col.grid = "grey",
           xlim=c(min_x,max_x),
           ylim=c(min_y,max_y),
           zlim=c(0,max_z)
           )
s3d$points3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei,
              z = combinePts$elevation,
              pch = 20,
              col= colours)

legend(s3d$xyz.convert(20, -0.5, -4),
       legend = c("forest cover random points", 
                  "research sites",
                  "literature sites", 
                  "inventory sites" 
                  ),
       col = c("#B1E5AD60", "#C9242C90", "#00CCCC", "#F6D65890"),
       pch = 20,
       cex = 0.7,
       inset = 0.00, xpd = TRUE, horiz = F)
```

```{r TemperateOceanicForest, echo=FALSE, message=FALSE, warning=FALSE}

#Look at 3d scatterplot
gezType <-'Temperate oceanic forest'
# ppPts<-dplyr::filter(pts30, id == 'pts' & gez == gezType )
combinePts<-dplyr::filter(combine, gez == gezType )
colourCombo <- c("#B1E5AD30", "#00CCCC", "#F6D65880", "#C9242C85")
# Pale Green: #B1E5AD30 pts
# Red: #C9242C research sites
# Aqua: "#00FFFF80"literature sites
# Yellow: #F6D65890 inventory sites
colours<-colourCombo[as.numeric(combinePts$type)]
s3d <- scatterplot3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei, 
              z = combinePts$elevation,
              pch = "",
              grid = F,
              box = T,
              # angle = 70,
              # color = colours,
              main = gezType,
              xlab = "Mean annual Temperature(°C)",
              ylab = "Water availability index",
              zlab = "Elevation (100m)",
              xlim=c(min_x,max_x),
              ylim=c(min_y,max_y),
              zlim=c(0,max_z)
              )

addgrids3d(x = combinePts$meanTemp, 
           y = combinePts$meanSpei, 
           z = combinePts$elevation,
           grid = c("xy", "xz", "yz"),
           # col.grid = "grey",
           xlim=c(min_x,max_x),
           ylim=c(min_y,max_y),
           zlim=c(0,max_z)
           )
s3d$points3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei,
              z = combinePts$elevation,
              pch = 20,
              col= colours)

legend(s3d$xyz.convert(20, -0.5, -4),
       legend = c("forest cover random points", 
                  "research sites",
                  "literature sites", 
                  "inventory sites" 
                  ),
       col = c("#B1E5AD60", "#C9242C90", "#00CCCC", "#F6D65890"),
       pch = 20,
       cex = 0.7,
       inset = 0.00, xpd = TRUE, horiz = F)

```

```{r SubtropicalMountainSystem, echo=FALSE, message=FALSE, warning=FALSE}

#Look at 3d scatterplot
gezType <-'Subtropical mountain system'
# ppPts<-dplyr::filter(pts30, id == 'pts' & gez == gezType )
combinePts<-dplyr::filter(combine, gez == gezType )
colours<-colourCombo[as.numeric(combinePts$type)]
s3d <- scatterplot3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei, 
              z = combinePts$elevation,
              pch = "",
              grid = F,
              box = T,
              # angle = 70,
              # color = colours,
              main = gezType,
              xlab = "Mean annual Temperature(°C)",
              ylab = "Water availability index",
              zlab = "Elevation (100m)",
              xlim=c(min_x,max_x),
              ylim=c(min_y,max_y),
              zlim=c(0,max_z)
              )

addgrids3d(x = combinePts$meanTemp, 
           y = combinePts$meanSpei, 
           z = combinePts$elevation,
           grid = c("xy", "xz", "yz"),
           # col.grid = "grey",
           xlim=c(min_x,max_x),
           ylim=c(min_y,max_y),
           zlim=c(0,max_z)
           )
s3d$points3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei,
              z = combinePts$elevation,
              pch = 20,
              col= colours)

legend(s3d$xyz.convert(20, -0.5, -4),
       legend = c("forest cover random points", 
                  "research sites",
                  "literature sites", 
                  "inventory sites" 
                  ),
       col = c("#B1E5AD60", "#C9242C90", "#00FFFF80", "#F6D65890"),
       pch = 20,
       cex = 0.7,
       inset = 0.00, xpd = TRUE, horiz = F)

```

```{r SubtropicalDryForest, echo=FALSE, message=FALSE, warning=FALSE}

#Look at 3d scatterplot
gezType <-'Subtropical dry forest'
# ppPts<-dplyr::filter(pts30, id == 'pts' & gez == gezType )

combinePts<-dplyr::filter(combine, gez == gezType & elevation >0 )
colours<-colourCombo[as.numeric(combinePts$type)]
s3d <- scatterplot3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei, 
              z = combinePts$elevation,
              pch = "",
              grid = F,
              box = T,
              # angle = 70,
              # color = colours,
              main = gezType,
              xlab = "Mean annual Temperature(°C)",
              ylab = "Water availability index",
              zlab = "Elevation (100m)",
              xlim=c(min_x,max_x),
              ylim=c(min_y,max_y),
              zlim=c(0,max_z)
              )

addgrids3d(x = combinePts$meanTemp, 
           y = combinePts$meanSpei, 
           z = combinePts$elevation,
           grid = c("xy", "xz", "yz"),
           # col.grid = "grey",
           xlim=c(min_x,max_x),
           ylim=c(min_y,max_y),
           zlim=c(0,max_z)
           )
s3d$points3d(x = combinePts$meanTemp, 
              y = combinePts$meanSpei,
              z = combinePts$elevation,
              pch = 20,
              col= colours)

legend(s3d$xyz.convert(20, -0.5, -4),
       legend = c("forest cover random points", 
                  "research sites",
                  "literature sites", 
                  "inventory sites" 
                  ),
       col = c("#B1E5AD60", "#C9242C90", "#00FFFF80", "#F6D65890"),
       pch = 20,
       cex = 0.7,
       inset = 0.00, xpd = TRUE, horiz = F)

```

